Identifying and correcting for misspecifications\\in GWAS summary statistics for polygenic scores

### 2022-02-04 paper/paper-misspec.tex
extend discu

In the real data applications hereinafter, we choose to use the first correction, ``sqrt\_info'', which is simple because it is equivalent to post-processing PGS effects by multiplying them by $\sqrt{\text{INFO}}$. [TODO: TRY THEM ALL AGAIN? AND META-ANALYZE (BRCA AND T1D)?]

@@

We have investigated misspecifications in GWAS summary statistics, focusing on the impact of sample size heterogeneity and imputation quality.
Previously, we proposed a quality control (QC) based on comparing standard deviations (of genotypes) inferred from GWAS summary statistics with the ones computed from a reference panel \cite[]{prive2020ldpred2}.
Here we show that we can refine this QC by deriving the latter directly from the reported allele frequencies in the GWAS summary statistics, and by correcting the former using the GWAS imputation INFO scores (e.g.\ see figure \ref{fig:brca_onco_qc}).
Using this refined QC, we are able to identify a potential issue with how principal components were derived in a set of breast cancer summary statistics. Fortunately, this has practically no effect on the predictive performance of the derived polygenic scores.
Additional QC can also be performed, e.g.\ comparing reported GWAS allele frequencies with the ones from the LD reference panel, e.g.\ to detect genotyping or allelic errors.
We do perform this additional QC as part of ``qc2'' here. 
One can also run other QC tools such as DENTIST \cite[]{chen2021improved}, and also infer ancestry proportions from summary statistics to make sure these are matching with the LD reference used \cite[]{prive2021using}.

@@

[TODO: PRS-CS REGU + SMALL LD REF UNMATCHED WITH ANCESTRY + LD BLOCKS]
[LD BLOCKS MAINLY FOR SCALA BUT SHOW THAT USEFUL FOR ROBUSTNESS]

@@

\subsection{LDpred2-low-h2 and LDpred2-auto-rob}

>>>>

In the real data applications hereinafter, we choose to use the first correction, ``sqrt\_info'', which is simple because it is equivalent to post-processing PGS effects by multiplying them by $\sqrt{\text{INFO}}$.

@@

We have investigated misspecifications in GWAS summary statistics, focusing particularly on the impact of sample size heterogeneity and imputation quality, and the application to polygenic scores (PGS) methods.
Previously, we proposed a quality control (QC) based on comparing standard deviations (of genotypes) inferred from GWAS summary statistics with the ones computed from a reference panel \cite[]{prive2020ldpred2}.
Here we show that we can refine this QC by deriving the latter directly from the reported allele frequencies in the GWAS summary statistics, and by correcting the former using the GWAS imputation INFO scores (e.g.\ see figure \ref{fig:brca_onco_qc}).
Using this refined QC, we are able to identify a potential issue with how principal components were derived in a set of breast cancer summary statistics. Fortunately, this has practically no effect on the predictive performance of the derived polygenic scores.
Additional QC can also be performed, e.g.\ comparing reported GWAS allele frequencies with the ones from the LD reference panel, e.g.\ to detect genotyping errors or allele inversions.
We do perform this additional QC as part of ``qc2'' here. 
One can also run other QC tools such as DENTIST \cite[]{chen2021improved}, and also infer ancestry proportions from summary statistics to make sure these are matching with the LD reference used \cite[]{prive2021using}.

@@

Anyway, we recommend to adopt these changes, i.e.\ performing the (refined) QC proposed here, defining LD blocks in the LD matrix, and using more regularization when needed (i.e.\ also testing smaller values for the heritability parameter in LDpred2-grid, and using the two new parameters in LDpred2-auto (Section \ref{sec:robust}). 
Note that LD blocks are widely used already by several methods (e.g.\ lassosum and PRS-CS) because they allow for processing smaller matrices at once \cite[]{mak2017polygenic,ge2019polygenic}; here we have shown that these blocks are also useful to make methods more robust.
PRS-CS is one of the most robust PGS methods out there; for example, it can use the (small) 1000 Genomes data as LD reference \cite[]{ge2019polygenic}, and can even use an European LD reference panel with multi-ancestry GWAS summary statistics \cite[]{wang2021global}.
We think this is made possible by the use of a (very) strong regularization in PRS-CS ($\phi^{-1} \psi_j^{-1} \ge 1$, which would approximately corresponds to using $s=0.5$ in lassosum, $\delta = 1$ in lassosum2, and $\texttt{shrink\_corr}=0.5$ in LDpred2-auto).
Using enough regularization is good for robustness, but note that using too much regularization can also damage predictive performance, and should not be a replacement for a proper QC and choice of ancestry-matched LD [TODO: A BIT AGGRESSIVE?].
As a future direction, we are interested in using ancestry-matched LD matrices to improve polygenic prediction in all ancestries.
Also it would be useful to investigate the impact of the QC and adjustments we propose here on other (non-PGS) methods, such as the ones for inference, where robustness is likely to be very important as well.

@@

\subsection{LDpred2-low-h2 and LDpred2-auto-rob}\label{sec:robust}
